### 2007-01-10 Documentation/appendix.tex
Part of manual updated
*** empty log message ***

@@

>>>>

{\bf NEEDS TO BE UPDATED [WW] -> PH? ADD SECTION ON CONDOR/CLUSTER SUPPORT?}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Flux normalization in \GLOBES }
\index{norm}{Normalization of fluxes}
\label{app:flux}

A common issue with \GLOBES\ is confusion about the proper units for
the input flux files for use in \AEDL\ experiment descriptions. Source of
the confusion is an undocumented factor 5.2 in \GLOBES\ versions older
than 3.0 the fluxes are multiplied with (see below). In Version 3.0 and 
higher, the alternative flux environment {\tt nuflux} is provided, 
which does not contain this factor. The following material is
based on the old environment {\tt flux}. For the use of {\tt nuflux},
replace the factor 5.20034 by unity.

\section*{Historical problem}

One problem for the design of \AEDL\ was initially that meaningful
units for flux data strongly depend on the given type of experiment,
but also on relatively arbitrary decisions. For accelerator beams
based on pion decay, one frequently defines the beam luminosity in
{\it protons on target (pot)} since this number has a one-to-one
correspondence with the number of neutrinos produced. Another sensible
unit could be {\it megawatt on target (MW)}, again this number is directly
correlated with the number of neutrinos and moreover there is a unique
relation to {\it pot} for a given accelerator. Of course,
what matters is the integrated luminosity. In some  cases the
neutrino flux is given per $10^7\,\mathrm{s}$. However,
most experiments will run for several years, hence also this number
has to enter somewhere. For neutrino factories the proper number is
useful muon decays per unit time, and for reactor experiments it is
the thermal power of the reactor {\it asf}. This demonstrates that it is
reasonable to keep the flux definition flexible.

\section*{Implementation in \GLOBES }

In understanding how one still can figure out what the correct units
are for each case, it is a good starting point to look at what
\GLOBES\ does with the input files. The cross section in the file is given as
differential cross section divided by energy $x=\sigma/E$, and the flux file
gives $f$. The differential number of events per GeV $n$ as computed in
\GLOBES\ without oscillation and efficiencies is given by
\begin{eqnarray}
n &=& 5.20034\times x\times E\times f\times \nonumber \\
&&\mathtt{@norm}\times\mathtt{@power}\times\mathtt{@stored\_muons}\times\mathtt{@time}\times\mathtt{\$target\_mass}\times(\mathtt{\$baseline})^{-2} \nonumber
\end{eqnarray}
Note that $5.20034$ is a undocumented fudge factor!

It is the sole responsibility of the author of the \AEDL\ file and its
supporting files, to ensure that the result makes sense. In
principle, it is possible to divide, for example, {\tt @time}  by
$\pi$ and fix that by redefining the flux file by multiplying
it with $\pi$. Modifications like that have happened in the past
and still happen, and many of them are not properly commented.

\section*{Writing \AEDL\ files}

The task is to choose the value of  {\tt @norm} such that all the
variables in the \AEDL\ file have the proper units, \eg , {\tt
  @time} has proper unit years.

\GLOBES\ assumes that the cross
section $x$ is given in $10^{-38}\,\mathrm{cm}^2$ and that all fluxes are
given at a distance of $1\,\mathrm{km}$. In addition, it assumes that
the number of target nuclei $\tau$ (or protons or whatever applies to the
given cross section) per unit target mass $m_u$ (which usually is $kt$)
 are properly accounted for.\footnote{Note that the cross sections which are delivered with
  \GLOBES\ always are per nucleon.}  Assuming that in the flux file
the data is given as number of neutrinos per unit area $A$ and
energy bin of width $\Delta E$ at a distance $L$ from the source, one
obtains
\begin{eqnarray}
\mathtt{@norm}=\frac{1}{5.20034}\left(\frac{\mathrm{GeV}}{\Delta
      E}\right)
\left(\frac{\mathrm{cm}^2}{A}\right)\left(\frac{L}{\mathrm{km}}\right)^2\left(\frac{\tau}{m_u}\right)\times10^{-38}\times\left(\frac{\mathcal{L}_u}{\mathcal{L}}\right)
\end{eqnarray}
where $\mathcal{L}$ absorbs all factors in the flux file related to
the integrated luminosity, and $\mathcal{L}_u$ is the unit chosen for
it.  The concept of integrated luminosity is
nicely described in the \GLOBES\ manual in \Sec~\ref{sec:source}.
A little example illustrates this concept:
The flux is given for $10^{21}\, \mathrm{pot}\,\mathrm{y}^{-1}$ of
$10\,\mathrm{GeV}$ protons, thus a good choice for the units $\mathcal{L}_u$
is $\mathrm{MW}\,\mathrm{y}^{-1}$, which means that
$\mathcal{L}/\mathcal{L}_u$ is given by (assuming a $10^7\,\mathrm{s}$
year)
\begin{equation}
\frac{\mathcal{L}}{\mathcal{L}_u}=\frac{10\,\mathrm{GeV}\,10^{21}\,\mathrm{pot}\,\mathrm{y}}{10^7\,\mathrm{s}}\times(\mathrm{MW}\,\mathrm{y}^{-1})^{-1}=0.160218\ldots
\end{equation}

### 2007-01-10 Documentation/part1.tex
*** empty log message ***

\begin{table}[tpb]
\begin{center}
\begin{tabular}{p{3.8cm}p{3.8cm}p{7cm}}
\hline
Function & Purpose & Parameters \ra\ Result \\
\hline
\multicolumn{3}{l}{{\bf Systematics only:}} \\
\GLB{ChiSys} & $\chi^2$ with systematics only  & ({\tt glb\_params in, int exp, int rule}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projections onto axes:}} \\
\GLB{ChiTheta13} & Projection onto $\theta_{13}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDelta} & Projection onto $\deltacp$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiTheta23} & Projection onto $\theta_{23}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDm31} & Projection onto $\ldm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDm21} & Projection onto $\sdm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto plane:}} \\
\GLB{ChiTheta13Delta} & Projection onto $\theta_{13}$-$\deltacp$-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto any hyper-plane:}} \\
\GLB{ChiNP} & Projection onto any $n$-dimensional hyper-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \newline
Needs \GLB{SetProjection} before! \\[0.2cm]
\multicolumn{3}{l}{{\bf Localization of degeneracies:}} \\
\GLB{ChiAll} & (Local) Minimization over all parameters  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\
\hline
\end{tabular}
\end{center}
\caption{\label{tab:stdfunctions} \index{norm}{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value with systematics only or systematics and correlations. The parameters {\tt rule} and {\tt exp}
can either be \GLBC{GLB\_ALL} for all initialized experiment or the
experiment number ($0$ to \GLB{\_num\_of\_exps}-1) for a specific experiment. The format of \GLB{\_params} is discussed in detail in \Chapt~\ref{chapt:gettingstarted}. Note that all functions but {\tt glbChiSys} are using minimizers which have to be initialized with \GLB{SetInputErrors} and \GLB{SetCentralValues} first.}
\end{table}

@@

Let us now discuss a simple example.  Typical applications for user-defined systematics are reactor neutrino experiments, where systematics play a crucial role. For a setup with near and far detectors,
the formula for $\chi^2$ is
\begin{align}
  \chi^2 =& \sum_{i=1}^{\textrm{\# of bins}} \sum_{d = N,F}2
     \left( O_{d,i} - (1 + a_R + a_d) T_{d,i} +
       (1 + a_R + a_d) T_{d,i} \log \frac{(1 + a_R + a_d) T_{d,i}}{O_{d,i}}  \right) \nonumber\\
         & \ \ + \frac{a_R^2}{\sigma_R^2} + \frac{a_N^2}{\sigma_N^2} + \frac{a_F^2}{\sigma_F^2}
\label{equ:chi2reactor}
\end{align}
where $O_{N,i}$ and $O_{F,i}$ are the event rates for the $i$-th bin in the near and
far detector, calculated for the assumed ``true'' values of the oscillation parameters.
$T_{d,i}$ are the event rates for the parameter values that are currently being tested.
$a_R$, $a_N$ and $a_F$ parametrize the small uncertainties in the reactor flux and the fiducial
mass of the two detectors. In this example their central values are assumed to be zero,
while their standard deviations are $\sigma_R$, $\sigma_N$ and $\sigma_F$.
%
The first line of \equ{chi2reactor} is the standard $\chi^2$ expression for
the Poisson distribution, while the terms in the second line are penalties
for deviations of the systematics parameters from their central values. 
%
We use two AEDL files for this experiment, one for the far detector and one for the
near detector.
The corresponding rule definition ($\nu_e$ disappearance) in the far detector AEDL-file is
\begin{quote}
{\tt
rule(\#rule0)< \\
\hspace*{0.5cm}        @signal     = 1.0@\#nu\_e\_disappearance\_CC \\
\hspace*{0.5cm}        @background = 0.0@\#nu\_e\_disappearance\_CC   /* No background */ \\
\hspace*{0.5cm} @energy\_window = 0.0015 : 0.01 \\
\\
\hspace*{0.5cm}        @sys\_off\_function = "chiNoSysSpectrum" \\
\hspace*{0.5cm}        @sys\_off\_errors   = \{ \} \\
\hspace*{0.5cm}        @sys\_on\_function  = "chiDCNorm" \\
\hspace*{0.5cm}        @sys\_on\_errors    = \{ 0.02,    0.006,        0.006,       0.005,     0.005   \} \\
\hspace*{0.5cm}        \mbox{ // \{ Flux, Fid. mass FD, Fid. mass ND, Energy FD, Energy ND \} } \\
> 
}
\end{quote}
In this case, the systematics is called ``{\tt chiDCNorm}'' for systematics on, whereas systematics
off is computed with spectral information, but without systematical errors.
For the near detector AEDL-file, we have instead
\begin{quote}
{\tt
rule(\#rule0)< \\
\hspace*{0.5cm}        @signal     = 1.0@\#nu\_e\_disappearance\_CC \\
\hspace*{0.5cm}        @background = 0.0@\#nu\_e\_disappearance\_CC   /* No background */ \\
\hspace*{0.5cm} @energy\_window = 0.0015 : 0.01 \\
\\
\hspace*{0.5cm}        @sys\_off\_function = "chiNoSysSpectrum" \\
\hspace*{0.5cm}        @sys\_off\_errors   = \{ \} \\
\hspace*{0.5cm}        @sys\_on\_function  = "chiZero" \\
\hspace*{0.5cm}        @sys\_on\_errors    = \{ \} \\
> 
}
\end{quote}
In this case, the systematics {\tt chiZero} is used for systematics on, which means that there will
be no active $\chi^2$ calculation in this rule. With this definition, the user-defined systematics
will only be called once for the far detector. However, the rates from the near detector will be 
passively provided for the common $\chi^2$ function.
You can find the corresponding files {\tt dchooz-near.glb} and {\tt dchooz-far.glb}, you will need for {\tt example5.c}, in the example directory. 
%
We show the implementation of the $\chi^2$ function in the example on page~\pageref{ex:syst},
where we in addition include an uncorrelated energy calibration error. See \figu{reactor} for the result
of this example.
%
\begin{figure}[t]
\begin{center}
\includegraphics[width=10cm]{reactor}
\end{center}
\caption{\label{fig:reactor} The result from {\tt example5.c} (figure similar in \Ref~\cite{Huber:2006vr}). The fat curve corresponds to the systematics on page~\pageref{ex:syst}.}
\end{figure}

@@

\begin{table}[t!]
\begin{tabular}{p{4.2cm}p{5.5cm}p{5.1cm}}
\hline
Function & Purpose & Parameters $\rightarrow$ Result\\
\hline
\GLB{AllocProjection} & Allocate projection vector & {\tt ()} \\
\GLB{FreeProjection} & Free projection vector {\tt stale} & {\tt (glb\_projection stale)} \\
\GLB{DefineProjection} & Assign projection vector {\tt in} & {\tt (glb\_projection in, int theta12, int theta13, int theta23, int delta, int dms, int dma)} \\ 
\GLB{CopyProjection} & Copy vector {\tt source} to {\tt dest} & {\tt (const glb\_projection source, glb\_projection dest)} \\
\GLB{PrintProjection} & Print vector {\tt in} to file {\tt stream} & {\tt (FILE* stream, const glb\_projection in)} \\
\GLB{SetProjectionFlag} & Set flag for oscillation parameter {\tt which} in vector {\tt in} to value {\tt flag}. & {\tt (glb\_projection in, int flag, int which)} \\
\GLB{GetProjectionFlag} & Return flag for oscillation parameter {\tt which} in vector {\tt in}. & {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
 {\tt glbSetDensity\-ProjectionFlag} \GLBNS{SetDensityProjectionFlag} & Set flag for density parameter {\tt which} in vector {\tt in} to value {\tt flag}. & {\tt (glb\_projection in, int flag, int which)} \\
{\tt glbGetDensity\-ProjectionFlag} \GLBNS{GetDensityProjectionFlag} & Return flag for density parameter {\tt which} in vector {\tt in}.  & {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
\hline
\end{tabular}
\caption{\label{tab:defprojection}\index{norm}{Projection!type}
Different functions handling the
\GLB{\_projection} type. Flags are either \GLBC{GLB\_FIXED} or \GLBC{GLB\_FREE}. The (un-shown) return values of the {\tt Set-} and {\tt Define-} functions point either to the assigned vector if successful, or they are {\tt NULL} if unsuccessful.}
\end{table}

>>>>

\begin{table}[tpb]
\begin{center}
\begin{tabular}{p{3.8cm}p{3.8cm}p{7cm}}
\hline
Function & Purpose & Parameters \ra\ Result \\
\hline
\multicolumn{3}{l}{{\bf Systematics only:}} \\
\GLB{ChiSys} & $\chi^2$ with systematics only  & ({\tt glb\_params in, int exp, int rule}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projections onto axes:}} \\
\GLB{ChiTheta13} & Projection onto $\theta_{13}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDelta} & Projection onto $\deltacp$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiTheta23} & Projection onto $\theta_{23}$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDm31} & Projection onto $\ldm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.1cm]
\GLB{ChiDm21} & Projection onto $\sdm$-axis  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto plane:}} \\
\GLB{ChiTheta13Delta} & Projection onto $\theta_{13}$-$\deltacp$-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\[0.2cm]
\multicolumn{3}{l}{{\bf Projection onto any hyper-plane:}} \\
\GLB{ChiNP} & Projection onto any $n$-dimensional hyper-plane  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \newline
Needs \GLB{SetProjection} before! \\[0.2cm]
\multicolumn{3}{l}{{\bf Localization of degeneracies:}} \\
\GLB{ChiAll} & (Local) Minimization over all parameters  &  ({\tt glb\_params in, glb\_params out, int exp}) \ra\  {\tt double} $\chi^2$ \\
\hline
\end{tabular}
\end{center}
\mycaption{\label{tab:stdfunctions} \index{norm}{Standard functions (table)} The \GLOBES\ standard function to obtain a $\chi^2$-value with systematics only or systematics and correlations. The parameters {\tt rule} and {\tt exp}
can either be \GLBC{GLB\_ALL} for all initialized experiment or the
experiment number ($0$ to \GLB{\_num\_of\_exps}-1) for a specific experiment. The format of \GLB{\_params} is discussed in detail in \Chapt~\ref{chapt:gettingstarted}. Note that all functions but {\tt glbChiSys} are using minimizers which have to be initialized with \GLB{SetInputErrors} and \GLB{SetCentralValues} first.}
\end{table}

@@

Let us now discuss a simple example.  Typical applications for user-defined systematics are reactor neutrino experiments, where systematics play a crucial role. For a setup with near and far detectors,
the formula for $\chi^2$ is
\begin{align}
  \chi^2 =& \sum_{i=1}^{\textrm{\# of bins}} \sum_{d = N,F}2
     \left( O_{d,i} - (1 + a_R + a_d) T_{d,i} +
       (1 + a_R + a_d) T_{d,i} \log \frac{(1 + a_R + a_d) T_{d,i}}{O_{d,i}}  \right) \nonumber\\
         & \ \ + \frac{a_R^2}{\sigma_R^2} + \frac{a_N^2}{\sigma_N^2} + \frac{a_F^2}{\sigma_F^2}
\label{equ:chi2reactor}
\end{align}
where $O_{N,i}$ and $O_{F,i}$ are the event rates for the $i$-th bin in the near and
far detector, calculated for the assumed ``true'' values of the oscillation parameters.
$T_{d,i}$ are the event rates for the parameter values that are currently being tested.
$a_R$, $a_N$ and $a_F$ parametrize the small uncertainties in the reactor flux and the fiducial
mass of the two detectors. In this example their central values are assumed to be zero,
while their standard deviations are $\sigma_R$, $\sigma_N$ and $\sigma_F$.
%
The first line of \equ{chi2reactor} is the standard $\chi^2$ expression for
the Poisson distribution, while the terms in the second line are penalties
for deviations of the systematics parameters from their central values. 
%
We use two AEDL files for this experiment, one for the far detector and one for the
near detector.
The corresponding rule definition ($\nu_e$ disappearance) in the far detector AEDL-file is
\begin{quote}
{\tt
rule(\#rule0)< \\
\hspace*{0.5cm}        @signal     = 1.0@\#nu\_e\_disappearance\_CC \\
\hspace*{0.5cm}        @background = 0.0@\#nu\_e\_disappearance\_CC   /* No background */ \\
\hspace*{0.5cm} @energy\_window = 0.0015 : 0.01 \\
\\
\hspace*{0.5cm}        @sys\_off\_function = "chiNoSysSpectrum" \\
\hspace*{0.5cm}        @sys\_off\_errors   = \{ \} \\
\hspace*{0.5cm}        @sys\_on\_function  = "chiDCNorm" \\
\hspace*{0.5cm}        @sys\_on\_errors    = \{ 0.02,    0.006,        0.006,       0.005,     0.005   \} \\
\hspace*{0.5cm}        \mbox{ // \{ Flux, Fid. mass FD, Fid. mass ND, Energy FD, Energy ND \} } \\
> 
}
\end{quote}
In this case, the systematics is called ``{\tt chiDCNorm}'' for systematics on, whereas systematics
off is computed with spectral information, but without systematical errors.
For the near detector AEDL-file, we have instead
\begin{quote}
{\tt
rule(\#rule0)< \\
\hspace*{0.5cm}        @signal     = 1.0@\#nu\_e\_disappearance\_CC \\
\hspace*{0.5cm}        @background = 0.0@\#nu\_e\_disappearance\_CC   /* No background */ \\
\hspace*{0.5cm} @energy\_window = 0.0015 : 0.01 \\
\\
\hspace*{0.5cm}        @sys\_off\_function = "chiNoSysSpectrum" \\
\hspace*{0.5cm}        @sys\_off\_errors   = \{ \} \\
\hspace*{0.5cm}        @sys\_on\_function  = "chiZero" \\
\hspace*{0.5cm}        @sys\_on\_errors    = \{ \} \\
> 
}
\end{quote}
In this case, the systematics {\tt chiZero} is used for systematics on, which means that there will
be no active $\chi^2$ calculation in this rule. With this definition, the user-defined systematics
will only be called once for the far detector. However, the rates from the near detector will be 
passively provided for the common $\chi^2$ function.
You can find the corresponding files {\tt dchooz-near.glb} and {\tt dchooz-far.glb}, you will need for {\tt example5.c}, in the example directory. 
%
We show the implementation of the $\chi^2$ function in the example on page~\pageref{ex:syst},
where we in addition include an uncorrelated energy calibration error. See \figu{reactor} for the result
of this example.
%
\begin{figure}[t]
\begin{center}
\includegraphics[width=10cm]{reactor}
\end{center}
\mycaption{\label{fig:reactor} The result from {\tt example5.c} (figure similar in \Ref~\cite{Huber:2006vr}). The fat curve corresponds to the systematics on page~\pageref{ex:syst}.}
\end{figure}

@@

\begin{table}[t!]
\begin{tabular}{p{4.2cm}p{5.5cm}p{5.1cm}}
\hline
Function & Purpose & Parameters $\rightarrow$ Result\\
\hline
\GLB{AllocProjection} & Allocate projection vector & {\tt ()} \\
\GLB{FreeProjection} & Free projection vector {\tt stale} & {\tt (glb\_projection stale)} \\
\GLB{DefineProjection} & Assign projection vector {\tt in} & {\tt (glb\_projection in, int theta12, int theta13, int theta23, int delta, int dms, int dma)} \\ 
\GLB{CopyProjection} & Copy vector {\tt source} to {\tt dest} & {\tt (const glb\_projection source, glb\_projection dest)} \\
\GLB{PrintProjection} & Print vector {\tt in} to file {\tt stream} & {\tt (FILE* stream, const glb\_projection in)} \\
\GLB{SetProjectionFlag} & Set flag for oscillation parameter {\tt which} in vector {\tt in} to value {\tt flag}. & {\tt (glb\_projection in, int flag, int which)} \\
\GLB{GetProjectionFlag} & Return flag for oscillation parameter {\tt which} in vector {\tt in}. & {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
 {\tt glbSetDensity\-ProjectionFlag} \GLBNS{SetDensityProjectionFlag} & Set flag for density parameter {\tt which} in vector {\tt in} to value {\tt flag}. & {\tt (glb\_projection in, int flag, int which)} \\
{\tt glbGetDensity\-ProjectionFlag} \GLBNS{GetDensityProjectionFlag} & Return flag for density parameter {\tt which} in vector {\tt in}.  & {\tt (const glb\_projection in, int which)} $\rightarrow$ {\tt int} flag \\
\hline
\end{tabular}
\mycaption{\label{tab:defprojection}\index{norm}{Projection!type}
Different functions handling the
\GLB{\_projection} type. Flags are either \GLBC{GLB\_FIXED} or \GLBC{GLB\_FREE}. The (un-shown) return values of the {\tt Set-} and {\tt Define-} functions point either to the assigned vector if successful, or they are {\tt NULL} if unsuccessful.}
\end{table}
